library(tidyverse)
library(lubridate)
library(data.table)
library(scales)

covid_data_tbl <- read_csv("https://opendata.ecdc.europa.eu/covid19/casedistribution/csv")


setDT(covid_data_tbl)
cumu_covid_tbl <- covid_data_tbl[order(-dateRep) & countriesAndTerritories %in% c("Germany", "United_Kingdom", "France", "Spain", "United_States_of_America") & year == 2020,.(dateRep, day, month, year, cases, cumulative_cases = rev(cumsum(rev(cases)))), by = countriesAndTerritories] %>%
  mutate(month = recode(.$month, "1" = "January", "2" = "February","3"="March", "4" = "April", "5"="May","6"="June","7"="July","8"="August","9"="September","10"="October","11"="November","12"="December")) %>%
  mutate(dateRep = as.Date(dateRep, "%d/%m/%Y")) %>% 
  mutate(countriesAndTerritories = as.factor(countriesAndTerritories)) 


```{r plot_4.1, fig.width=10, fig.height=7}
require("ggrepel")
cumu_covid_tbl %>%
  
  ggplot(aes(dateRep, cumulative_cases, color = countriesAndTerritories)) +
  geom_line(size = 0.5) +
  scale_x_date(date_breaks = "1 month", date_labels = "%B" )  +
  scale_y_continuous(labels =  number_format(scale = 1e-6, suffix = " M")) +
 
    labs(
      title = "COVID-19 confirmed cases worldwide",
      subtitle = "As of 12/05/2020, Europe has more Cases then the USA",
      tag = "Challenge 1",
      x = "Year 2020",
      y = "Comulative Cases",
      color = "Country:"
    ) + 
theme(
      axis.text.x = element_text(angle = 45, hjust = 1),
      legend.position = "bottom",
      plot.title = element_text(face = "bold"),
      plot.tag = element_text(face = "bold"),
      plot.tag.position = "bottom"
    )
```


library(mapdata)
world <- map_data("world")

mortality_rate_tbl <- covid_data_tbl[,.(population = mean(popData2019), sum_deaths = sum(deaths)),by = countriesAndTerritories][,mortality_rate :=(sum_deaths/population)] %>% 
  mutate(across(countriesAndTerritories, str_replace_all, "_", " ")) %>%
  mutate(countriesAndTerritories = case_when(
    countriesAndTerritories == "United Kingdom" ~ "UK",
    countriesAndTerritories == "United States of America" ~ "USA",
    countriesAndTerritories == "Czechia" ~ "Czech Republic",
    TRUE ~ countriesAndTerritories
  ))

### 2. Data visualisation
```{r plot_4.2, fig.width=10, fig.height=7}

mortality_rate_tbl %>% 
  ggplot(aes(map_id= countriesAndTerritories)) +
  geom_map(aes(fill = mortality_rate), map = world ) +
  expand_limits(x = world$long, y = world$lat) +
  scale_fill_continuous(low    = "yellow", high = "red",
                        labels =  percent_format(),
                        limits=c(0,0.0015),
                        breaks=seq(0,0.0015,by=0.0003))+
  # Formatting
    labs(
      title = "Confirmed COVID-19 deaths relative to the size of the population",
      subtitle = "More then 1.2 Million confirmed COVID-19 death worldwide",
      caption = "Date: 06.12.2020",
      tag = "Challenge 2",
      fill = "Mortality Rate"
    ) + 
  theme_dark() +
theme(
      axis.title=element_blank(),
      axis.text=element_blank(),
      axis.ticks=element_blank(),
      legend.position = "right",
      plot.title = element_text(face = "bold"),
      plot.tag = element_text(face = "bold"),
      plot.tag.position = "bottom"
    )
```